---
title: "Experiment 7"
date: 2023-11-06
output: html_notebook
---

#Packages 
```{r}
library(tidyverse)
library(janitor)
library(here)
library(openmeteo)
```

###Other Project Defaults
```{r}
theme_set(theme_minimal())
```
# Read in Bikedata
```{r}
bikes <- here("data_raw", "202309-capitalbikeshare-tripdata.csv")

dfbikes <- read_csv(bikes) |> clean_names()
```
# Get the type of bike, type of rider, with start time, and ride_id, and rename start
```{r}
df2s = dfbikes |>
  select(rideable_type, member_casual, contains("start"), ride_id)|>
  mutate(start_stop = "start") |>
           rename(t = started_at, 
                  station_name = start_station_name,
         station_id = start_station_id,
         lat= start_lat,
         lng=start_lng
         )


```

```{r}
df2e=dfbikes %>% 
  select(ride_id,rideable_type,member_casual,
                    contains("end")) %>%
  mutate(start_stop="stop") %>%
  rename(t=ended_at,
         station_name=end_station_name,
         station_id=end_station_id,
         lat=end_lat,
         lng=end_lng)
```
```{r}
df2s = dfbikes |>
  select(rideable_type, member_casual, contains("start"), ride_id)|>
  mutate(start_stop = "start") |>
           rename(t = started_at, 
                  station_name = start_station_name,
         station_id = start_station_id,
         lat= start_lat,
         lng=start_lng
         )
```

```{r}
df2e=dfbikes %>% 
  select(ride_id,rideable_type,member_casual,
                    contains("end")) %>%
  mutate(start_stop="stop") %>%
  rename(t=ended_at,
         station_name=end_station_name,
         station_id=end_station_id,
         lat=end_lat,
         lng=end_lng)
```

```{r}
df2s=dfbikes %>% 
  select(rideable_type,member_casual,
                    contains("start"),ride_id) %>% 
  mutate(start_stop="start") %>%
  rename(t=started_at,
         station_name=start_station_name,
         station_id=start_station_id,
         lat=start_lat,
         lng=start_lng)

df2e=dfbikes %>% 
  select(ride_id,rideable_type,member_casual,
                    contains("end")) %>%
  mutate(start_stop="stop") %>%
  rename(t=ended_at,
         station_name=end_station_name,
         station_id=end_station_id,
         lat=end_lat,
         lng=end_lng)
 
df2=bind_rows(df2s,df2e) %>%
  arrange(t) %>%
  mutate(rider_delta=(start_stop=="start")*2-1) %>% #change in ridership 
  mutate(riders=cumsum(rider_delta)) %>%
  relocate(riders,.after=t)


```

#This depicts a full months view of bikeridership
```{r}
df2 %>% 
  ggplot(aes(t,riders)) +
  geom_line()
```
#This depicts a full months view of bikeridership in geompoint form
```{r}
df2 %>% 
  ggplot(aes(t,riders)) +
  geom_point()
```

# Constructing a Subsampled dataset of bikeriders

```{r}
df_s=df2 |> slice_head(n=1000)
```
# Depiction of riders by the amount of time they ride
```{r}
df_s %>% 
  ggplot(aes(t,riders)) +
  geom_point()
```
#
```{r}
df_e=df_s |>
  mutate(t_f=floor_date(t, "10 mins")) |>
  relocate(t_f, .after=t) |>
  slice_head(n=1, by= t_f)
```
# The visulaztion depicts the one above with the geompoints used, but this uses geomline instead
```{r}
df_e %>% 
  ggplot(aes(t,riders)) +
  geom_line()
```
# 
```{r}
df_r=df2 |>

  mutate(t_f=floor_date(t,"10 mins")) %>%

  relocate(t_f,.after=t) %>%

  slice_head(n=1,by=t_f)
```

```{r}
df_r %>% 
  ggplot(aes(t,riders)) +
  geom_line()
```
```{r}
p1=df2 %>% 
  filter(day(t)==18) %>%
  ggplot(aes(t,riders)) +
  geom_line() +
  ggtitle("Riders on 18Sep")

p1+
  geom_line(data=df_r %>% filter(day(t)==18),
  color="red")
```
```{r}
df_r=df2 |>
  mutate(t_f=floor_date(t,"1 mins")) %>%
  relocate(t_f,.after=t) %>%
  slice_head(n=1,by=t_f)
```
# Selection of weather history within DC from sept 1st to sept 30, 2023, factoring in appearent tempreature, wind speed over 10 meters, and pereceptation
```{r}
df_w=weather_history("Washington",
                    start = "2023-09-01",
                    end = "2023-09-30",
                    hourly = c("apparent_temperature",
                               "wind_speed_10m",
                               "precipitation")
)
```

```{r}
df_w %>% 
  ggplot(aes(datetime, hourly_precipitation)) +
  geom_line()
```

```{r}
df_r %>% 
  ggplot(aes(t,riders)) +
  geom_line()
```
```{r}
df_j=df_s %>% left_join(df_w,
                        by=join_by(closest(t>=datetime)))
```
# Joining and relocating tables/columns like datetime to be after t
```{r}
df_j=df_s %>% 
  left_join(df_w,by=join_by(closest(t>=datetime)))  %>%
  relocate(datetime, .after=t)
 
head(df_j)
```
# Thus is needed to convert the system time to that of the New York Time, due to DC and New York Having the same time
```{r}
df_j$t[1:5]
df_j$datetime[1:5]

df_jb <- force_tz(df_j, tzone = "America/New_York")
df_jb
df_sb <-force_tz(df_s, tzone = "America/New_York")
df_sb

df_jb$t[1:5]
df_jb$datetime[1:5]

df_sb$t[1:5]
df_sb$datetime[1:5]

```